---
title: "Data Science Homework Problem"
output: word_document
editor_options: 
  chunk_output_type: console
---
 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      tidy = 'styler', 
                      comment = NA,
                      message = FALSE,
                      warning = FALSE)
```

This assignment is intended to quickly assess general data science skills such as data munging (manipulation,
reshaping, joining) and exploratory analysis.

Completion of the tasks should only require about an hour of your time. The goal is to showcase general skills, so it is not necessary to provide lengthy answers or polished results. An answer that simply addresses each question will suffice. If it takes you longer than an hour, that is okay, but we don’t want to take up too
much of your time with this exercise.

The dataset that will be used for this exercise is related to American Major League Baseball statistics. This
is intended to simulate scenarios where a data scientist must work with data that may be in an unfamiliar
domain.

There are two data files. The main data file is `hits.csv`. Each record of this dataset pertains to an “at-bat”,
where a ball is thrown to a batter and the batter has a number of tries to hit the ball. The dataset has records
for 5,000 at-bats in the 2020 baseball season that resulted in a hit. The target in this dataset is whether or
not the batter hits a “home run”, which is hitting the ball over the outfield wall. Variables recorded for each
at-bat include the batter’s name and team, the pitcher’s name and team, the speed of the pitch, the angle
the ball comes off the bat, etc. While no data dictionary is provided, the variables of interest in this dataset
will be described in the exercises.
Please provide the source code for your solutions. You can submit your solution however you would like – an
R script with comments, an R Markdown document, a Word document, etc.


```{r R packages, include=TRUE, echo=FALSE}
packages_needed <- c("tidyverse", "lubridate", "knitr", "kableExtra",  "htmltools", "mime")

if (!require(install.load)) {
  install.packages("install.load")
}

install.load::install_load(packages_needed)

theme_set(theme_bw())
```

## Read in `hits.csv`

You should get something that looks like this:

```{r, echo=FALSE}
hits <- read_csv("hits.csv")

hits
```


# Explore the number of hits over time**

Using the `game_date` variable, create a visualization that shows the number of hits over time. Do you
notice anything strange about the distribution? If so, what might you do to further investigate? (Actual investigation not necessary - just a quick answer will suffice).


## Create a univariate exploratory plot for launch speed

The continuous variable `launch_speed` is the speed at which the ball leaves the bat after it is hit. Provide a simple univariate exploratory visualization of this variable.

## Create a univariate exploratory plot for the bearing of the hit ball

The bearing of the hit ball is the direction that the ball travels with respect to the baseball field, one of either “right”, “center” or “left” field. Create a summary visualization of this categorical variable.

## Visually compare the distribution of launch speed based on whether the hit was a “home run” or not

The variable `is_home_run` is $1$ if the hit ball goes over the outfield wall (called a “home run”), or $0$ if the hit ball does not go over the wall. Create a visualization that compares the distribution of launch speed for hits that are home runs and hits that are not.


## Visually compare the distribution of bearing based on whether the hit was a “home run” or not

Similar to (5), create a visualization that compares the distribution of “bearing” and whether or not the hit is a home run.

## Calculate a summary of the percentage of home runs for each batter

For each batter, denoted by the variable batter_name, calculate the percentage of home runs that they hit. Which batter has the highest home run percentage? Explain for this batter whether you would trust the computed statistic or not.
## Read in `park_dimensions.csv`

The file **park_dimensions.csv** contains information about every baseball park in American Major League
Baseball, including how far away the left, center, and right field walls of the baseball park are from the batter, and the height of the walls. It is thought that augmenting our hits dataset with some of this information could give us more insight into whether a hitter will hit a home run. The data should look something like this:

```{r, echo=FALSE}
park <- read_csv("park_dimensions.csv")

park
```

## Pivot the park dimensions from wide format to long format

The variables LF_Dim, CF_Dim, RF_Dim provide the distance to the left field, center field, and right field walls, in feet. Create a new dataset from these variables that is pivoted from wide to long format, which should result in something that looks like this:

```{r, echo=FALSE}
clean_park <- park %>% pivot_longer(cols = c(LF_Dim, CF_Dim, RF_Dim), names_to = "bearing", values_to = "wall_distance") %>% mutate(bearing = case_when(bearing == "LF_Dim" ~ "left", bearing == "CF_Dim" ~"center", TRUE ~ "right")) %>% select(park, bearing, wall_distance)

clean_park
```

## Join the wall distance with the hits data

For each record in the hits dataset, we have the baseball park ID stored in park and the bearing of the hit ball stored in bearing. Join your results from (9) to the hits dataset using these variables so the resulting hits dataset has a new column indicating the distance to the wall for the given baseball park and bearing of each hit.
